---
title: "Create and Preprocess AML-ETO SO"
author: "TLusardi"
date: "11/02/2020"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
params:
  local: TRUE
---

### Set up libraries and directories

```{r "setup_libs", include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(data.table)
library(ggplot2)
library(knitr)
library(networkD3)
library(patchwork)
library(dplyr)

runs2process = c("dmso", "ava", "ory", "combo")
#importNew <- "clust"
#importDate <- "2020-10-26"

# ADT Antibodies
antibodies <- c(CD45RA = "ab_CD45RA", CD123 = "ab_CD123", CD10 = "ab_CD10", CD90 = "ab_CD90",
                CD34 = "ab_CD34", CD38 = "ab_CD38", CD366 = "ab_CD366", CD99 = "ab_CD99")
adt.ab2gene <- c('ab-CD45RA' = "PTPRC", 'ab-CD123' = "IL3RA", 'ab-CD10' = "MME", 'ab-CD90' = "THY1",
                 'ab-CD34' = "CD34", 'ab-CD38' = "CD38", 'ab-CD366' = "HAVCR2", 'ab-CD99' = "CD99")
```

```{r "setup_vars", include=FALSE}
directory = list(raw = "/Users/lusardi/Documents/CEDAR/Projects/3.ASXL_mutant/1.AML_ETO/analysis/data/raw",
                 rda = "/Users/lusardi/Documents/CEDAR/Projects/3.ASXL_mutant/1.AML_ETO/analysis/data/rda")

# If raw exists, continue, otherwise, exit.
if (!dir.exists(directory$raw)) {
  message(sprintf("ERROR: source directory does not exist %s", directory$raw))
  knitr::knit_exit()
}

# If rda does not exist, create it
if (!dir.exists(directory$rda)) {
  dir.create(directory$rda)
  if (!dir.exists(directory$rda)) {
    message(sprintf("ERROR: could not create directory %s", directory$rds))
    knitr::knit_exit()
  }
}

# Create a list of variables used to create the SO
ab_gene <- c(ab_CD45RA = "PTPRTC", ab_CD123 = "IL3RA", ab_CD10 = "MME", ab_CD90 = "THY1",
             ab_CD34 = "CD34", ab_CD38 = "CD38", ab_CD366 = "HAVCR2", ab_CD99 = "CD99")
settings.ls <- list()

```

### Read in data by run

Customize here as required per cellranger output naming convention
* Note that "_" is reserved in Seurat (used as delimiter between an object key and a feature name)
  +  https://github.com/satijalab/seurat/issues/3219
  

```{r, "readraw", echo=FALSE}
# Get a list of directories with data
runs <- list.dirs(path = directory$raw, full.names = FALSE, recursive = FALSE)
# Select the runs of interest
runs <- runs[grepl("amleto", runs)]
select_index <- 2
names(runs) <- tolower(tstrsplit(runs, "_")[[select_index]])
# Consistent name w/TB
names(runs)[names(runs) == "avaory"] <- "combo"

# Read in by run
expts.ls <- list()
for (myrun in runs2process) { 
  mydata = Read10X(data.dir = sprintf("%s/%s/outs/filtered_feature_bc_matrix", directory$raw, runs[[myrun]]))

  # Gene Expression matrix
  gex <- mydata$`Gene Expression`
  
  # Antibody matrix 
  #  Replace _ with - to suppress seurat error
  adt <- mydata$`Antibody Capture`[,]
  adt_abs <- rownames(adt)
  rownames(adt) <- gsub("_", "-", rownames(adt))
  
  updateADT <- FALSE
  if (updateADT) {
    # Fix the antibody names 
    # ** THIS CAN BE DONE at the CellRanger phase - just put the NAME in. I like ab_<name> so it can be distinguished from the rna
    myfeatures <- fread(sprintf("%s/%s/outs/feature_reference.csv", directory$raw, runs[[myrun]]))
    myfeatures[, matrix.rowName := paste(name, .I-1, sep = ".")]
    myfeatures[matrix.rowName == paste(name, 0, sep = "."), matrix.rowName := name]
    myfeatures[, antibody := tstrsplit(id, "-")[[6]]]
    myfeatures[, abID := paste("ab", antibody, sep = "_") ]
    myfeatures[, abGene := adt.ab2gene[abID]]
    adt.rownames <- myfeatures$abID
    names(adt.rownames) <- myfeatures$matrix.rowName
  
    # Update ADT row names
    if (sum(rownames(adt) %in% myfeatures$matrix.rowName) > 0) {
      rownames(adt) <- adt.rownames[rownames(adt)]
    }
  }
  
  # Update column names 
  colnames(gex) <- paste(myrun, gsub("-1", "", colnames(gex)), sep = ".")
  colnames(adt) <- paste(myrun, gsub("-1", "", colnames(adt)), sep = ".")
  
  # Confirm that the same cells are present in the same order for each matrix
  if (all.equal(colnames(adt), colnames(gex))) {
    print("Column order consistent - antibody and gene expression")
  } else {
    print("Column order inconsistent among antibody and gene expression - need to fix it.")
  }

  # Consider the counts per gene
  print("***********************************************")
  # Summarize Features
  print(sprintf("CellRanger Output %s: %i features in %i cells", myrun,
                dim(gex)[1], dim(gex)[2]))
  minFeat <- min(Matrix::rowSums(gex))
  print(sprintf("Feature counts range %i to %i", minFeat, max(Matrix::rowSums(gex))))
  print(sprintf("Features with %i counts: %i", minFeat,
                sum(Matrix::rowSums(gex) == minFeat)))
  print(sprintf("Features with 1-3 counts: %i", 
                sum(Matrix::rowSums(gex) > 0 & Matrix::rowSums(gex) < 4)))
  minCells <- min(Matrix::colSums(gex))
  print(sprintf("Cell feature counts range %i to %i", minCells,
                max(Matrix::colSums(gex))))
  print(sprintf("Cells with %i counts: %i", minCells,
                sum(Matrix::colSums(gex) == minCells)))

  # Summarize antibody counts
  print(sprintf("CellRanger Antibody Output %s: %i features in %i cells", myrun,
                dim(adt)[1], dim(adt)[2]))
  minAb <- min(Matrix::rowSums(adt))
  print(sprintf("Antibody counts range %i to %i", minAb, max(Matrix::rowSums(adt))))
  print(sprintf("Antibody with %i counts: %i", minAb,
                sum(Matrix::rowSums(adt) == minAb)))
  minAbCells <- min(Matrix::colSums(adt))
  print(sprintf("Cell antibody counts range %i to %i", minAbCells,
                max(Matrix::colSums(adt))))
  print(sprintf("Cells with %i counts: %i", minAbCells,
                sum(Matrix::colSums(adt) == minAbCells)))

  # Create Seurat Object
  # Set thresholds for cell, feature inclusion
  min.cells = 3
  min.features = 200
  settings.ls[["step_CreateSeuratObject"]][["min.cells"]] <- min.cells
  settings.ls[["step_CreateSeuratObject"]][["min.features"]] <- min.features

  # Initialize the Seurat object with the raw (non-normalized data).
  myso <- CreateSeuratObject(counts = gex, project = myrun,
                             min.cells = min.cells, min.features = min.features)
  
  print(sprintf("%s: %i cells in ADT matrix that are not present in gene expression matrix", myrun, sum(!(colnames(adt) %in% colnames(gex)))))
  print(sprintf("%s: %i cells in gene expression matrix that are not present in ADT matrix", myrun, sum(!(colnames(gex) %in% colnames(adt)))))
  
  myso[["ADT"]] <- CreateAssayObject(counts = adt[, colnames(myso)])
  
  if (updateADT) {
    myso@misc$adt.features <- myfeatures
  }
  
  # Define adt - gene relationship
  myso@misc$ab_gene <- ab_gene
  
  expts.ls[[myrun]] <- myso
}
```

### Take a quick look at ADT concordance

```{r, echo = FALSE}
for (myrun in names(expts.ls)) { 
  myso <- expts.ls[[myrun]]
  for (myab in rownames(myso@assays$ADT)) {
    if (updateADT) {
      myfeatures <- myso@misc$adt.features
      mygene <- myfeatures[abID == gsub("-", "_", myab), abGene]
    } else  {
      mygene <- adt.ab2gene[myab]
    }
    mygene <- gsub("ab_", "", mygene)
    if (!(mygene %in% rownames(myso))) {
      print(sprintf("%s:  %s RNA present in fewer than %i cells in %s object", myab, mygene, min.cells, myrun))
    } else {
      plot(myso@assays$RNA[mygene, ], myso@assays$ADT[myab,], log = "y", pch = 20, xlab = mygene, ylab = myab, main = myrun)
    }
  }
}
```

### Summary Plots
Summarize data to make filtering decisions

```{r, "summaryPlot", echo=FALSE}
for (myrun in names(expts.ls)) {
  myso <- expts.ls[[myrun]]
  
  # Quantify % mitochondrial features
  myso[["percent.mt"]] <- PercentageFeatureSet(myso, pattern = "^MT-")
  plot(VlnPlot(myso, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), pt.size = 0.5, ncol = 3) )
  
  # Plot density function of percent.mt
  # Consider how much to include/exclude
  colors = c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')
  probs = c(0.99, 0.95, 0.9, 0.85, 0.80, 0.75, 0.70)
  cut_opts = quantile(myso@meta.data$percent.mt, probs = probs)
  p1 <- plot(density(myso@meta.data$percent.mt), main = sprintf("Distribution of %% Mitochondrial Cells %s", myrun))
  for (i in 1:length(cut_opts)) {
    p1 <- abline(v = cut_opts[i], col = colors[i])
  }
  legend_text = paste(probs, round(cut_opts, 1), sep = " - ")
  p1 <- legend(x = "topright", legend = legend_text, col = colors[1:length(probs)], pch = 16)
  
  # Plot density function of nFeature_RNA
  # Consider how much to include/exclude
  colors = c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')
  probs = c(0.99, 0.95, 0.9, 0.85, 0.15, 0.10, 0.05, 0.01)
  cut_opts = quantile(myso@meta.data$nFeature_RNA, probs = probs)
  p1 <- plot(density(myso@meta.data$nFeature_RNA), main = sprintf("Distribution of nFeature_RNA Cells %s", myrun))
  for (i in 1:length(cut_opts)) {
    p1 <- abline(v = cut_opts[i], col = colors[i])
  }
  legend_text = paste(probs, round(cut_opts, 1), sep = " - ")
  p1 <- legend(x = "topright", legend = legend_text, col = colors[1:length(probs)], pch = 16)
  expts.ls[[myrun]] <- myso
  
  # Plot density function of nCount_RNA
  # Consider how much to include/exclude
  colors = c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')
  probs = c(0.99, 0.95, 0.9, 0.85, 0.15, 0.10, 0.05, 0.01)
  cut_opts = quantile(myso@meta.data$nCount_RNA, probs = probs)
  p1 <- plot(density(myso@meta.data$nCount_RNA), main = sprintf("Distribution of nCount_RNA Cells %s", myrun))
  for (i in 1:length(cut_opts)) {
    p1 <- abline(v = cut_opts[i], col = colors[i])
  }
  legend_text = paste(probs, round(cut_opts, 1), sep = " - ")
  p1 <- legend(x = "topright", legend = legend_text, col = colors[1:length(probs)], pch = 16)
  expts.ls[[myrun]] <- myso

}
```
  
### Preprocess Runs
Based on summary plots, filter out cells with excess mitochondrial RNA, low/high counts
Assign Cell Cycle Scores

* Note that the cell cycle scores are influenced by the population of cells (before and after filtering give different results)  
- A casual look showed that all of the scores are different, differences are all different (so it's not just a shift), and a small number of Phase assignments are different (~10% from an n=1 off the cuff analysis))


```{r, "preprocess", echo=FALSE}
s.genes <- cc.genes.updated.2019$s.genes
g2m.genes <- cc.genes.updated.2019$g2m.genes

preproc.expts.ls <- list()
for (myrun in names(expts.ls)) { 
  myso <- expts.ls[[myrun]]

  # Filter high mitochondria cells
  mitocut = 12
  settings.ls[["step_filter"]][["mitocut"]] <- mitocut
  nmitofilt <- sum(myso@meta.data$percent.mt > mitocut)
  print(sprintf("%s:  Filter cells > %i%% mitochondria: %i/%i (%.1f%%)", myrun, mitocut, nmitofilt, ncol(myso), nmitofilt/ncol(myso)*100))
  
  # Filter high and low RNA count cells
  minCount = 200
  maxCount = 25000
  settings.ls[["step_filter"]][["minCount"]] <- minCount
  settings.ls[["step_filter"]][["maxCount"]] <- maxCount
  nfeatfilt <- sum(myso@meta.data$nCount_RNA < minCount | myso@meta.data$nCount_RNA > maxCount)
  print(sprintf("%s:  Filter cells with %i > nCount > %i: %i/%i (%.1f%%)",
                myrun, minCount, maxCount, nfeatfilt, ncol(myso), nfeatfilt/ncol(myso)*100))
  
  # Add settings info to the object under "misc"
  myso@misc$settings <- settings.ls
  
  myso <- subset(myso, subset = nCount_RNA > minCount & nCount_RNA < maxCount & percent.mt < mitocut)
  print(sprintf("Filtered %s:  %i cells, %i features", myrun, dim(myso)[2], dim(myso)[1]))
  
  # Assign cell cycle scores
  myso <- CellCycleScoring(myso, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
  
  # Per Seurat "Alternate Workflow" to reduce loss of relevant cell cycle information (eg, during hematopoiesis)
  myso$CC.Difference <- myso$S.Score - myso$G2M.Score
  
  preproc.expts.ls[[myrun]] <- myso
}
```

### Normalize & Scale 
Quick look to assess regression needs

```{r, "regressPeek", echo=FALSE}
vars2regress <- list(none = "none", mito = "percent.mt",
                     CC = "Phase", CC.mito = c("Phase", "percent.mt"),
                     Diff = "CC.Difference", Diff.mito = c("CC.Difference", "percent.mt"))
  
for (myrun in runs2process[1]) {
  for (myregress in names(vars2regress)) {
    myso <- preproc.expts.ls[[myrun]]
    
    # Normalize counts
    myso <- NormalizeData(myso)
  
    # Identify variable genes
    genecount <- nrow(myso)
    myso <- FindVariableFeatures(myso, selection.method = "vst", nfeatures = 2000, verbose = FALSE)
  
    # Scale Counts
    if (myregress == "none") {
      myso <- ScaleData(myso)
    } else {
      myso <- ScaleData(myso, vars.to.regress = vars2regress[[myregress]])
    }
  
    # Perform PCA
    myso <- RunPCA(myso)
  
    plot1 <- (DimPlot(myso, reduction = "pca", group.by = "Phase", split.by = "Phase") +
           plot_annotation(title = sprintf("%s - %s regression", myrun, myregress)))
    plot2 <- (FeaturePlot(myso, reduction = "pca", features = "percent.mt", split.by = "Phase") +
           plot_annotation(title = sprintf("%s - %s regression", myrun, myregress)))
    plot(plot1 + plot2)
  }
}
```

### Assign Cell-Cycle Scores
So - filter first, then assign cell cycle scores

```{r, "assignCC", echo=FALSE}
# Plot some cell cycle genes
for (myrun in names(preproc.expts.ls)) {
  myso <- preproc.expts.ls[[myrun]]
  plot(RidgePlot(myso, features = c("NASP", "USP1", "TUBB4B", "HMGB2"), group.by = "Phase", ncol = 2))
  print(sprintf("Sum the counts in s-phase genes across %i cells", ncol(myso)))
  print(Matrix::rowSums(myso@assays$RNA@counts[rownames(myso) %in% s.genes, ]))
  print(sprintf("Sum the counts in g2m-phase genes across %i cells", ncol(myso)))
  print(Matrix::rowSums(myso@assays$RNA@counts[rownames(myso) %in% g2m.genes, ]))
}
```

### Save Data
If you add filtering, make sure that you adjust the file saving accordingly

```{r, echo=FALSE}
file2save <- sprintf("aml_eto.preprocSO.nofilt.%s.rds", Sys.Date())
print(sprintf("Saving preprocessed data in individual objects in %s", file2save))
saveRDS(preproc.expts.ls, file = paste(directory$rda, file2save, sep = "/"))
```

```{r, echo=FALSE}
sessionInfo()
```

