---
title: "AML-ETO SO - integrate, sct, cluster"
author: "TLusardi"
date: "11/05/2020"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
params:
  local: TRUE
---

Integrate objects   
* Use all features so can pass object to P3

Assign Cell cycle scores
* Can compare assignment to individual run assignments

Run SCTransform on individual objects   
* regress - Cell cycle
  + noregress
  + regress.CC - regress on cell cycle - probably not the best option as it removes any cell cycle information that may be important for differentiating cells
  + regress.diff - regress on the difference in cell cycle values; per Seurat notations, it preserves cell cycle information relevant to development
* regress - other features
  + nCounts_RNA + percent.mt
  
Cluster & Plot!

Save each object to a separate file

importDate - select the creation date for the desired object
  
### Set up libraries and directories

```{r "setup_libs", include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(data.table)
library(ggplot2)
library(knitr)
library(networkD3)
library(patchwork)
library(dplyr)

runs2process = c("dmso", "ava", "ory", "combo")
importDate <- "2020-11-03"
```

```{r "setup_vars", include=FALSE}
directory = list(raw = "/Users/lusardi/Documents/CEDAR/Projects/3.ASXL_mutant/1.AML_ETO/analysis/data/raw",
                 rda = "/Users/lusardi/Documents/CEDAR/Projects/3.ASXL_mutant/1.AML_ETO/analysis/data/rda")

expts.ls <- readRDS(sprintf("%s/aml_eto.preprocSO.nofilt.%s.rds", directory$rda, importDate))
```

### Preview Clustering

```{r, echo=FALSE}

```

### Plot Counts Comparisons

```{r, echo=FALSE}
plotSet = list(dmso = list(col = "#e41a1c", lty = 1),
               ava = list(col = "#377eb8", lty = 1),
               ory = list(col = "#4daf4a", lty = 1),
               combo = list(col = "#adaf4a", lty = 1))
runs <- names(expts.ls)

plotFeatures <- c("nCount_RNA", "nFeature_RNA")
for (myfeature in plotFeatures) {
  # Set up Density Plot
  maxDens <- 0
  maxCts <- 0
  for (myrun in names(expts.ls)) {
    mydens <- density(expts.ls[[myrun]]@meta.data[[myfeature]])
    maxDens <- max(maxDens, mydens$y)
    maxPct <- 1
    maxCts <- max(maxCts, expts.ls[[myrun]]@meta.data[[myfeature]])
  }

  # Create a plot pane
  plot(NULL, xlim=c(0, maxCts), ylim = c(0, maxDens), 
       xlab = sprintf("%s", myfeature),
       ylab = "Density", las = 1,
       main = sprintf("Distribution of %s per Cell by Experiment", myfeature))
  abline(h = 0, col = "grey80")

  # Add a legend
  legend(x="topright", legend = runs, lwd = 2,
         col = unlist(plotSet)[paste(runs, "col", sep = ".")],
         lty = as.numeric(unlist(plotSet)[paste(runs, "lty", sep = ".")]))

  # Plot each density line
  for (myrun in names(expts.ls)) {
    lines(density(expts.ls[[myrun]]@meta.data[[myfeature]]), col = plotSet[[myrun]]$col,
          lwd = 2)
  }
}

temp <- merge(x = expts.ls$dmso, y = expts.ls[c("ava", "ory", "combo")])
temp <- merge(x = expts.ls$ava, y = expts.ls[c("dmso", "ory", "combo")])
temp <- merge(x = expts.ls$ory, y = expts.ls[c("dmso", "ava", "combo")])

all.equal(expts.ls$dmso@assays$RNA@counts, temp@assays$RNA@counts[rownames(temp) %in% rownames(expts.ls$dmso), grepl("dmso", colnames(temp))])
all.equal(expts.ls$ava@assays$RNA@counts, temp@assays$RNA@counts[rownames(temp) %in% rownames(expts.ls$ava), grepl("ava", colnames(temp))])
all.equal(expts.ls$ory@assays$RNA@counts, temp@assays$RNA@counts[rownames(temp) %in% rownames(expts.ls$ory), grepl("ory", colnames(temp))])
all.equal(expts.ls$combo@assays$RNA@counts, temp@assays$RNA@counts[rownames(temp) %in% rownames(expts.ls$combo), grepl("combo", colnames(temp))])

plot(x = expts.ls$ory@assays$RNA@counts, y = temp@assays$RNA@counts[rownames(temp) %in% rownames(expts.ls$combo), grepl("combo", colnames(temp))] )
```

### Integrate the objects

```{r, "integrate", echo=FALSE}
# Needed to avoid error in getGlobalsandPackages 
#options(future.globals.maxSize= 3530*1024^2)

expts.integrated.ls <- list()
for (myregress_feat in c("no", "yes")) {
  for (myregress in c("SCT", "SCT.CC", "SCT.Diff")) {
    my.features <- SelectIntegrationFeatures(object.list = expts.ls[[myfilt]], nfeatures = 3000, assay = rep(myregress, 4))
    myiso <- PrepSCTIntegration(object.list = expts.ls[[myfilt]], anchor.features = my.features, verbose = FALSE, assay = rep(myregress, 4))
    my.anchors <- FindIntegrationAnchors(object.list = myiso, normalization.method = "SCT", reference = 1, scale = FALSE,
                                         anchor.features = my.features, verbose = TRUE)
    integratedName <- paste("Integrated", myregress, sep = "_")
    myintegrated <- IntegrateData(anchorset = my.anchors, normalization.method = "SCT", verbose = FALSE, new.assay.name = integratedName)
    expts.integrated.ls[[myfilt]][[myregress]] <- myintegrated
  }
}
```

### Integrate


### Normalize & Scale
Use SCTransform with regression options noted

```{r, "runSCT", warning=FALSE, echo=FALSE}
regressions <- list(noregression = "NULL",
                    regress.CC = c("S.Score", "G2M.Score"),
                    regress.Diff = c("CC.Difference"))    

for (myfilt in names(expts.ls)) {
  for (myregress in names(regressions)) {
    for (myrun in names(expts.ls[[myfilt]])) {
      if (myregress == "noregression") {
        myassay <- "SCT"
        expts.ls[[myfilt]][[myrun]] <- SCTransform(expts.ls[[myfilt]][[myrun]], verbose = FALSE, new.assay.name = myassay)
      } else {
        myassay <- gsub("regress", "SCT", myregress)
        expts.ls[[myfilt]][[myrun]] <- SCTransform(expts.ls[[myfilt]][[myrun]], verbose = FALSE, new.assay.name = myassay, vars.to.regress = regressions[[myregress]])
      }
    
      # Calculate pca
      mypca <- gsub("SCT", "pca", myassay)
      mypcakey <- paste(gsub("/.", "", mypca), "_", sep = "")
      expts.ls[[myfilt]][[myrun]] <- RunPCA(expts.ls[[myfilt]][[myrun]], verbose = FALSE, assay = myassay,
                                  reduction.name = mypca, reduction.key = mypcakey)
      
      # Create umap
      myumap <- gsub("pca", "umap", mypca)
      myumapkey <- paste(gsub("/.", "", myumap), "_", sep = "")
      expts.ls[[myfilt]][[myrun]] <- RunUMAP(expts.ls[[myfilt]][[myrun]], dims = 1:30, assay = myassay,
                                   reduction = mypca, reduction.name = myumap, reduction.key = myumapkey)
    } 
  }
}
```

### Plot UMAP and PCA

```{r, echo=FALSE}
for (myfilt in names(expts.ls)) {
  for (myrun in names(expts.ls[[myfilt]])) {
    for (myregress in names(regressions)) {
      myassay <- ifelse(myregress == "noregression", "SCT", gsub("regress", "SCT", myregress))
      mypca <- gsub("SCT", "pca", myassay)
      plot(DimPlot(expts.ls[[myfilt]][[myrun]], reduction = mypca)  + labs(title = sprintf("%s: %s", myrun, myregress)))
    
      myumap <- gsub("SCT", "umap", myassay)
      plot(DimPlot(expts.ls[[myfilt]][[myrun]], reduction = myumap)  + labs(title = sprintf("%s: %s", myrun, myregress)))
    }
  }
}
```

### Save the data

```{r, echo=FALSE}
# Save the file
for (myfilt in names(expts.ls)) {
  file2save <- sprintf("aml_eto.regressSO.%s.%s.rds", myfilt, Sys.Date())
  print(sprintf("%s objects:  Saving scaled/normalized/regressed data in individual objects in %s", myfilt, file2save))
  saveRDS(expts.ls[[myfilt]], file = paste(directory$rda, file2save, sep = "/"))
}
```

```{r, echo=FALSE}
sessionInfo()
```
