---
title: "Cluster AML-ETO SO"
author: "TLusardi"
date: "11/04/2020"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
params:
  local: TRUE
---

#### Operations   
* Cluster each object individually
* Crudely identify clusters in the DMSO object
* Transfer cluster identities to the ava, ory, and combo objects

##### Data Import Options - File naming
*importDate - select the creation date for the desired object
* Filtering - build in ability to process differently filtered objects in parallel
  + nofilt:  objects include all cells that pass min/max counts, max % mitochondria
  + filtered:  objects filtered to have similar numbers of cells, counts/cell  
  
##### Regression Options (in each object)
* regress - start from preprocessed objects
  + noregress
  + regress.CC - regress on cell cycle - probably not the best option as it removes any cell cycle information that may be important for differentiating cells
  + regress.diff - regress on the difference in cell cycle values; per Seurat notations, it preserves cell cycle information relevant to development
  
  
### Set up libraries and directories

```{r "setup_libs", include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(data.table)
library(ggplot2)
library(knitr)
library(networkD3)
library(patchwork)
library(dplyr)

filts2process = c("nofilt", "filtered")[1]
importDate <- "2020-11-03"
runs2process = c("dmso", "ava", "ory", "combo")
```

```{r "readObjects", include=FALSE}
directory = list(raw = "/Users/lusardi/Documents/CEDAR/Projects/3.ASXL_mutant/1.AML_ETO/analysis/data/raw",
                 rda = "/Users/lusardi/Documents/CEDAR/Projects/3.ASXL_mutant/1.AML_ETO/analysis/data/rda")

expts.ls <- list()
for (myfilt in filts2process) {
  expts.ls[[myfilt]] <- readRDS(sprintf("%s/aml_eto.regressSO.%s.%s.rds", directory$rda, myfilt, importDate))
}
```

### Cluster individual experiments

```{r, "iCluster", echo=FALSE, warning=FALSE, message=FALSE}
for (myfilt in filts2process) { 
  for (myrun in names(expts.ls[[myfilt]])) {
    myso <- expts.ls[[myfilt]][[myrun]]
    for (myregress in c("SCT", "SCT.CC", "SCT.Diff")) {
      mypca <- gsub("SCT", "pca", myregress)
      mygraph <- paste(myregress, "snn", sep = "_")
    
      myso <- FindNeighbors(myso, dims = 1:10, assay = myregress, reduction = mypca)
      myso <- FindClusters(myso, graph.name = mygraph, resolution = 0.8)
    }
    expts.ls[[myfilt]][[myrun]] <- myso
  }
}
```

### Plot Individual Clusters

```{r, "iplotCluster", echo=FALSE, warning=FALSE, message=FALSE}
for (myfilt in filts2process) {
  for (myrun in names(expts.ls[[myfilt]])) {
    for (myregress in c("SCT", "SCT.CC", "SCT.Diff")) {
      mypca <- gsub("SCT", "pca", myregress)
      mypcakey <- paste(gsub("/.", "", mypca), "_", sep = "")
      myumap <- gsub("SCT", "umap", myregress)
      myumapkey <- paste(gsub("/.", "", myumap), "_", sep = "")
      mygraph <- paste(myregress, "snn", sep = "_")
      myclusters <- paste(myregress, "snn_res.0.8", sep = "_")
      plot(DimPlot(expts.ls[[myfilt]][[myrun]], reduction = myumap, group.by = myclusters)
           + labs(title = sprintf("%s - %s: %s regression", myfilt, myrun, myregress)))
    }
  }
}
```

###Sankey Plots to illustrate clustering differences

```{r, "compSankey", echo=FALSE}
  
for (myfilt in filts2process) {
  for (myrun in names(expts.ls[[myfilt]])) {
    myso <- expts.ls[[myfilt]][[myrun]]
  
    # Create node and link files from meta.data
    mymeta <- data.table(myso@meta.data, keep.rownames = "CellID")
    mymeta[, SCT_clust := paste("SCT", SCT_snn_res.0.8, sep = "_")]
    mymeta[, SCT.CC_clust := paste("SCT.CC", SCT.CC_snn_res.0.8, sep = "_")]
    mymeta[, SCT.Diff_clust := paste("SCT.Diff", SCT.Diff_snn_res.0.8, sep = "_")]
    nodes <- data.table(Name = c(unique(mymeta$SCT_clust), unique(mymeta$SCT.CC_clust), unique(mymeta$SCT.Diff_clust)))
    nodes <- nodes[order(Name)]
    nodes[, nodeID := .I - 1]
    links12 <- mymeta[, .N, by = .(SCT_clust, SCT.Diff_clust)][order(SCT_clust)]
    colnames(links12) <- c("source", "target", "value")
    links23 <- mymeta[, .N, by = .(SCT.Diff_clust, SCT.CC_clust)][order(SCT.Diff_clust)]
    colnames(links23) <- c("source", "target", "value")
    links <- rbindlist(list(links12, links23))
    links <- merge(x = links, y = nodes, by.x = "source", by.y = "Name", all.x = TRUE, all.y = FALSE)
    colnames(links)[colnames(links) == "nodeID"] <- "SourceID"
    links <- merge(x = links, y = nodes, by.x = "target", by.y = "Name", all.x = TRUE, all.y = FALSE)
    colnames(links)[colnames(links) == "nodeID"] <- "TargetID"
  
    print(sankeyNetwork(Links = links, Nodes = nodes, Source = "SourceID", Target = "TargetID",
                        Value = "value", NodeID = "Name", fontSize = 12))
  
    cols2transfer <- c("SCT_clust", "SCT.CC_clust", "SCT.Diff_clust")
    if (all.equal(mymeta$CellID, colnames(myso))) {
      for (mycol in cols2transfer) {
        myso[[mycol]] <- mymeta[[mycol]]
      }
      expts.ls[[myfilt]][[myrun]] <- myso
    } else {
      print(sprintf("%s: Need to order the metadata to replace in SO", myrun))
    }
  }
}
```

### Identify dmso Clusters (loosely!)
CD34/SOX4/ERG are stem cell genes one end of the dog bone  
CD14, ITGAX, LYZ are mature monocyte genes at the other end of the dog bone

Note: Will need to redo this in filtered objects

```{r, "clusterID", echo=FALSE, fig.height=6, fig.width=6}
idMarkers <- c("CD34", "SOX4", "ERG", "MYB", "CEBPD", "GATA2", "CD14", "ITGAX", "LYZ")
idMarkers <- c("SOX4", "CEBPD", "CD14")
for (myfilt in filts2process) {
  for (myrun in "dmso") {
    for (myregress in c("SCT", "SCT.CC", "SCT.Diff")) {
      myumap <- gsub("SCT", "umap", myregress)
      spliton <- paste(myregress, "snn_res.0.8", sep = "_")
      DimPlot(expts.ls[[myfilt]][[myrun]], reduction = myumap, group.by = "orig.ident")
      plot(DimPlot(expts.ls[[myfilt]][[myrun]], reduction = myumap, group.by = spliton, label = TRUE, label.size = 6) +
             plot_annotation(title = sprintf("%s - %s:  Individual Object Cluster ID, %s regression", myfilt, myrun, myregress)))
      plot(FeaturePlot(expts.ls[[myfilt]][[myrun]], features = idMarkers, reduction = myumap, slot = "scale.data", pt.size = 0.25) +
             plot_annotation(title = sprintf("%s - %s:  Marker Genes, %s regression", myfilt, myrun, myregress)))
    }
  }
}
```

### Assign DMSO cluster names

```{r, "clusterName", echo=FALSE, fig.height=6, fig.width=6}
# Cluster Labels
dmsoLabels.ls <- list(nofilt = c(SCT_0 = "late_1", SCT_1 = "early_1_2", SCT_2 = "early_1b", SCT_3 = "early_2b",
                                 SCT_4 = "mid_1", SCT_5 = "lone_1", SCT_6 = "mid_2", SCT_7 = "mid_3b",
                                 SCT_8 = "lone_2", SCT_9 = "lone_mid", SCT_10 = "mid_3a",
                                 SCT.Diff_0 = "late_1", SCT.Diff_1 = "early_1", SCT.Diff_2 = "early_2",
                                 SCT.Diff_3 = "mid_3", SCT.Diff_4 = "mid_1", SCT.Diff_5 = "mid_2",
                                 SCT.Diff_6 = "lone_1", SCT.Diff_7 = "lone_2", SCT.Diff_8 = "lone_mid",
                                 SCT.CC_0 = "late_1", SCT.CC_1 = "early_1", SCT.CC_2 = "early_2",
                                 SCT.CC_3 = "mid_2_3", SCT.CC_4 = "mid_1", SCT.CC_5 = "lone_1",
                                 SCT.CC_6 = "lone_2", SCT.CC_7 = "mid_3b", SCT.CC_8 = "lone_mid") )

dmsoLevels.ls <- list(nofilt = c("early_1", "early_2", "early_1_2", "early_1b", "early_2b",
                                 "mid_1", "mid_2", "mid_3", "mid_2_3", "mid_3a" , "mid_3b",
                                 "late_1", "lone_mid", "lone_1", "lone_2"))

for (myfilt in filts2process) {
  for (myrun in "dmso") {
    dmsoLabels <- dmsoLabels.ls[[myfilt]]
    dmsoLevels <- dmsoLevels.ls[[myfilt]]
    myso <- expts.ls[[myfilt]][[myrun]]
    for (myregress in c("SCT", "SCT.CC", "SCT.Diff")) {
      # Add in the cluster label
      myclust <- paste(myregress, "clust", sep = "_")
      mylabelname <- paste(myclust, "label", sep = "_")
      myso[[mylabelname]] <- factor(dmsoLabels[unlist(myso[[myclust]])], levels = dmsoLevels)
    
      # Plot with updated names
      myumap <- gsub("SCT", "umap", myregress)
      spliton <- paste(myregress, "snn_res.0.8", sep = "_")
      plot(DimPlot(myso, reduction = myumap, group.by = mylabelname, label = TRUE, label.size = 6) +
             plot_annotation(title = sprintf("%s - %s:  Individual Object Cluster Names, %s regression", myfilt, myrun, myregress)))
#      plot(FeaturePlot(expts.ls[[myfilt]][[myrun]], features = idMarkers, reduction = myumap, slot = "scale.data", pt.size = 0.25))
    }
    
    # Update expts.ls
    expts.ls[[myfilt]][[myrun]] <- myso
  }
}
```

### Transfer Cluster Labels

```{r, "transferLabels", echo=FALSE}
for (myfilt in filts2process) {
  for (myregress in c("SCT", "SCT.CC", "SCT.Diff")) {
    for (myrun in c("ava", "ory", "combo")) {
      myfeatures <- intersect(VariableFeatures(expts.ls[[myfilt]]$dmso), VariableFeatures(expts.ls[[myfilt]][[myrun]]))
      my.anchors <- FindTransferAnchors(reference = expts.ls[[myfilt]]$dmso, query = expts.ls[[myfilt]][[myrun]], dims = 1:30, 
                                        reference.assay = myregress, query.assay = myregress,
                                        features = myfeatures,
                                        normalization.method = "SCT")
      myrefcol <- paste(myregress, "clust_label", sep = "_")
      my.predictions <- TransferData(anchorset = my.anchors,
                                     refdata = expts.ls[[myfilt]]$dmso@meta.data[,myrefcol],
                                     dims = 1:30)
      colnames(my.predictions) <- paste(myregress, colnames(my.predictions), sep = "_")
      expts.ls[[myfilt]][[myrun]] <- AddMetaData(expts.ls[[myfilt]][[myrun]], metadata = my.predictions)
    }
  }
}

for (myfilt in filts2process) {
  for (myrun in c("ava", "ory", "combo")) {
    myso <- expts.ls[[myfilt]][[myrun]]
    for (myregress in c("SCT", "SCT.CC", "SCT.Diff")) {
      # Add in the cluster label
      myclust <- paste(myregress, "clust", sep = "_")
      mylabelname <- paste(myregress, "predicted.id", sep = "_")
      
      # Plot with updated names
      myumap <- gsub("SCT", "umap", myregress)
      spliton <- paste(myregress, "snn_res.0.8", sep = "_")
      plot(DimPlot(myso, reduction = myumap, group.by = spliton, label = TRUE, label.size = 6, repel = TRUE) +
             labs(title = sprintf("%s: %s - %s, cluster identification", myfilt, myregress, myrun)))
      plot(DimPlot(myso, reduction = myumap, group.by = mylabelname, label = TRUE, label.size = 6, repel = TRUE) +
             labs(title = sprintf("%s: %s - %s, labels transferred from DMSO", myfilt, myregress, myrun)))
    }
  }
}
```

### Compare transferred vs. native clusters

```{r, "compTxNativeClust", echo=FALSE}
for (myfilt in filts2process) {
  for (myrun in c("ava", "ory", "combo")) {
    myso <- expts.ls[[myfilt]][[myrun]]
    mymeta <- data.table(myso@meta.data, keep.rownames = "CellID")
    for (myregress in c("SCT", "SCT.CC", "SCT.Diff")) {
      # Add in the cluster label
      myclust <- paste(myregress, "clust", sep = "_")
      mylabelname <- paste(myregress, "predicted.id", sep = "_")
      
      # Node and Link for Sankey
      nodes <- data.table(Name = c(unique(mymeta[[myclust]]), unique(mymeta[[mylabelname]])))
      nodes <- nodes[order(Name)]
      nodes[, nodeID := .I - 1]
      links <- mymeta[, .N, by = mget(c(myclust, mylabelname))][order(get(mylabelname))]
      colnames(links) <- c("source", "target", "value")
      links <- merge(x = links, y = nodes, by.x = "source", by.y = "Name", all.x = TRUE, all.y = FALSE)
      colnames(links)[colnames(links) == "nodeID"] <- "SourceID"
      links <- merge(x = links, y = nodes, by.x = "target", by.y = "Name", all.x = TRUE, all.y = FALSE)
      colnames(links)[colnames(links) == "nodeID"] <- "TargetID"
      print(sankeyNetwork(Links = links, Nodes = nodes, Source = "SourceID", Target = "TargetID",
                          Value = "value", NodeID = "Name", fontSize = 12))
    }
  }
}
```

### Save the data

```{r, echo=FALSE}
# Save the file
for (myfilt in names(expts.ls)) {
  file2save <- sprintf("aml_eto.indivClustSO.%s.%s.rds", myfilt, Sys.Date())
  print(sprintf("%s objects:  Saving scaled/normalized/regressed data in individual objects in %s", myfilt, file2save))
  saveRDS(expts.ls[[myfilt]], file = paste(directory$rda, file2save, sep = "/"))
}
```

```{r, echo=FALSE}
sessionInfo()
```


