---
title: "Create AML-ETO SO"
author: "TLusardi"
date: "10/17/2020"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
params:
  local: TRUE
---

#### Operations   
* Integrate objects by regression
* Cluster Integrated Object
* Compare clustering - integrated vs. inidvidual with transferred cluster identities

##### Data Import Options - File naming
*importDate - select the creation date for the desired object
* Filtering - build in ability to process differently filtered objects in parallel
  + nofilt:  objects include all cells that pass min/max counts, max % mitochondria
  + filtered:  objects filtered to have similar numbers of cells, counts/cell  
  
##### Regression Options (in each object)
* regress - start from preprocessed objects
  + noregress
  + regress.CC - regress on cell cycle - probably not the best option as it removes any cell cycle information that may be important for differentiating cells
  + regress.diff - regress on the difference in cell cycle values; per Seurat notations, it preserves cell cycle information relevant to development
  
### Set up libraries and directories

```{r "setup_libs", include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(data.table)
library(ggplot2)
library(knitr)
library(networkD3)
library(patchwork)
library(dplyr)
filts2process = c("nofilt", "filtered")[1]
importDate <- "2020-11-04"
runs2process = c("dmso", "ava", "ory", "combo")
```


```{r "readObjects", include=FALSE}
directory = list(raw = "/Users/lusardi/Documents/CEDAR/Projects/3.ASXL_mutant/1.AML_ETO/analysis/data/raw",
                 rda = "/Users/lusardi/Documents/CEDAR/Projects/3.ASXL_mutant/1.AML_ETO/analysis/data/rda")

expts.ls <- list()
for (myfilt in filts2process) {
  expts.ls[[myfilt]] <- readRDS(sprintf("%s/aml_eto.indivClustSO.%s.%s.rds", directory$rda, myfilt, importDate))
}
```

### Integrate the objects

Can probably collapse this into the expts.ls object... Don't want to validate the process right now.

```{r, "integrate", echo=FALSE}
# Needed to avoid error in getGlobalsandPackages 
options(future.globals.maxSize= 3530*1024^2)

expts.integrated.ls <- list()
for (myfilt in filts2process) {
  for (myregress in c("SCT", "SCT.CC", "SCT.Diff")) {
    my.features <- SelectIntegrationFeatures(object.list = expts.ls[[myfilt]], nfeatures = 3000, assay = rep(myregress, 4))
    myiso <- PrepSCTIntegration(object.list = expts.ls[[myfilt]], anchor.features = my.features, verbose = FALSE, assay = rep(myregress, 4))
    my.anchors <- FindIntegrationAnchors(object.list = myiso, normalization.method = "SCT", reference = 1, scale = FALSE,
                                         anchor.features = my.features, verbose = TRUE)
    integratedName <- paste("Integrated", myregress, sep = "_")
    myintegrated <- IntegrateData(anchorset = my.anchors, normalization.method = "SCT", verbose = FALSE, new.assay.name = integratedName)
    expts.integrated.ls[[myfilt]][[myregress]] <- myintegrated
  }
}
```

### Dim reduction on integrated objects

```{r, "integrated_dimreduct", warning=FALSE, echo=FALSE, fig.height=6, fig.width=6}
for (myfilt in filts2process) {
  for (myregress in names(expts.integrated.ls[[myfilt]])) {
    myso <- expts.integrated.ls[[myfilt]][[myregress]]
    myassay <- paste("Integrated", myregress, sep = "_")
    myreduction <- paste("pca", myassay, sep = "_")
    myumap <- gsub("pca", "umap", myreduction)
    myso <- RunPCA(myso, verbose = FALSE, assay = myassay, reduction.name = myreduction)
    myso <- RunUMAP(myso, dims = 1:30, assay = myassay, reduction = myreduction, reduction.name = myumap)
  
    # Create a factor for orig ident
    exptlevels <- c("dmso", "ava", "ory", "combo")
    myso$expt.fac <- factor(myso$orig.ident, levels = exptlevels)
  
    expts.integrated.ls[[myfilt]][[myregress]] <- myso
  }
}
```

### Plot Integrated results
use per-experiment clusters
  
```{r, "plot_integrated_dimreduct", warning=FALSE, echo=FALSE, fig.height=6, fig.width=6}
# This should ultimately be stored in so@misc from step 3.
dmsoLevels.ls <- list(nofilt = c("early_1", "early_2", "early_1_2", "early_1b", "early_2b",
                                 "mid_1", "mid_2", "mid_3", "mid_2_3", "mid_3a" , "mid_3b",
                                 "late_1", "lone_mid", "lone_1", "lone_2"))
idMarkers <- c("SOX4", "CEBPD", "CD14")

for (myfilt in filts2process) {
  dmsoLevels <- dmsoLevels.ls[[myfilt]]
  for (myregress in names(expts.integrated.ls[[myfilt]])) {
    myso <- expts.integrated.ls[[myfilt]][[myregress]]
    mymeta <- data.table(myso@meta.data, keep.rownames = "CellID")
  
    # Combine cluster labels from dmso and transferred labels (this should be up where the labels are transferred)
    mymeta[, txClusterID := get(paste(myregress, "predicted.id", sep = "_"))]
    mymeta[orig.ident == "dmso", txClusterID := get(paste(myregress, "clust_label", sep = "_"))]
    mymeta[, txClusterID.fac := factor(txClusterID, levels = dmsoLevels)]
    if (all.equal(colnames(myso), mymeta$CellID)) {
      myso[["txClusterID"]] <- mymeta$txClusterID
      myso[["txClusterID.fac"]] <- mymeta$txClusterID.fac
    } else {
      print("ERROR: Cluster Labels not transferred")
    }
  
    myassay <- paste("Integrated", myregress, sep = "_")
    myreduction <- paste("pca", myassay, sep = "_")
    myumap <- gsub("pca", "umap", myreduction)
    groupby <- paste(myregress, "snn_res.0.8", sep = "_")
  
    # Plot with per run x regression clusters
    myplot <- DimPlot(myso, split.by = "expt.fac", group.by = groupby, ncol = 2)
    plot(myplot + plot_annotation(title = "Cluster Labels from Individual Sample", subtitle = myregress))
  
    # Plot with transferred cluster labels
    myplot <- DimPlot(myso, split.by = "expt.fac", group.by = "txClusterID.fac", ncol = 2)
    plot(myplot + plot_annotation(title = "Cluster Labels transferred from DMSO", subtitle = myregress))
  
    myplot <- FeaturePlot(myso, features = idMarkers[1:3], split.by = "expt.fac",
                          reduction = myumap, slot = "scale.data", pt.size = 0.25)
    plot(myplot + plot_annotation(title = "Cluster Marker Genes", subtitle = myregress))
  
    # Update expts.integrated.ls with new meta.data
    expts.integrated.ls[[myfilt]][[myregress]] <- myso
  }
}
```

### Cluster Integrated objects
Not sure that this is necessary. Could be interesting, but not right now.

```{r, echo=FALSE}
for (myfilt in filts2process) {
  for (myregress in names(expts.integrated.ls[[myfilt]])) {
    myso <- expts.integrated.ls[[myfilt]][[myregress]]
    myassay <- paste("Integrated", myregress, sep = "_")
    myreduction <- paste("pca", myassay, sep = "_")
    myumap <- gsub("pca", "umap", myreduction)
  
    myso <- FindNeighbors(myso, reduction = myreduction, dims = 1:10, assay = myassay)
    mygraph <- paste(myassay, "snn", sep = "_")
    myso <- FindClusters(myso, graph.name = mygraph, resolution = 0.8)
  
    expts.integrated.ls[[myregress]] <- myso
  }
}
```

### Plot cluster data
Skip unless reclustering the integrated object

```{r, "plotClusters", echo=FALSE}
for (myfilt in filts2process) {
  for (myregress in names(expts.integrated.ls[[myfilt]])) {
    mygroup <- sprintf("Integrated_%s_snn_res.0.8", myregress)
    myassay <- paste("Integrated", myregress, sep = "_")
    myreduction <- paste("pca", myassay, sep = "_")
    myumap <- gsub("pca", "umap", myreduction)
    plot(DimPlot(expts.integrated.ls[[myregress]], split.by = "expt.fac", group.by = mygroup, ncol = 2))
    plot(FeaturePlot(expts.integrated.ls[[myregress]], features = idMarkers, reduction = myumap, slot = "scale.data", pt.size = 0.25))
  }
}
```

### Find Markers for Each Cluster

```{r, "FindMarkersIntegrated", echo=FALSE, fig.height = 10, fig.width=8}
for (myfilt in filts2process) {
  for (myregress in c("SCT.Diff")) {
    myso <- expts.integrated.ls[[myfilt]][[myregress]]
    Idents(myso) <- "txClusterID.fac"
    myassay <- paste("Integrated", myregress, sep = "_")
    mymarkers <- FindAllMarkers(myso, assay = myassay, min.pct = 0.25) 
    mydmsoMarkers <- FindAllMarkers(subset(myso, subset = orig.ident == "dmso"), assay = myassay, min.pct = 0.25)
  }
}

# Set Thresholds
lfc.min <- 1
pval.max <- 0.005
Idents(myso) <- "txClusterID.fac"

ClusterMarkers.dt <- data.table(mymarkers, keep.rownames = "GeneRow")
filtmarkers <- ClusterMarkers.dt[avg_logFC > lfc.min & p_val_adj < pval.max, ]
top10 <- filtmarkers %>% group_by(cluster) %>% top_n(n=10, wt = avg_logFC)

DmsoMarkers.dt <- data.table(mydmsoMarkers, keep.rownames = "GeneRow")
filtdmsomarkers <- DmsoMarkers.dt[avg_logFC > lfc.min & p_val_adj < pval.max, ]
top10dmso <- filtdmsomarkers %>% group_by(cluster) %>% top_n(n=10, wt = avg_logFC)

myplot <- DoHeatmap(myso, features = top10$gene) + NoLegend()
plot(myplot + plot_annotation(title = sprintf("%s, Integrated Object, adjp < %.3f", myregress, pval.max)))
myplot <- DoHeatmap(myso, features = top10dmso$gene) + NoLegend()
plot(myplot + plot_annotation(title = sprintf("%s, DMSO Object, adjp < %.3f", myregress, pval.max)))
```
### ID DE Genes - DMSO vs. Combo

```{r, "deCombo", echo=FALSE}
myso <- expts.integrated.ls$nofilt$SCT.Diff
clustDE.ls <- list()
Idents(myso) <- "orig.ident"
for (myrun in c("ava", "ory", "combo")) {
  runclust.ls <- list()
  allclust <- levels(myso@meta.data$txClusterID.fac)[levels(myso@meta.data$txClusterID.fac) %in% unique(myso@meta.data$txClusterID)]
  for (myclust in allclust) {
    mysubso <- subset(myso, subset = txClusterID.fac == myclust)
    runclust.ls[[myclust]] <- data.table(FindMarkers(mysubso, ident.1 = "dmso", ident.2 = myrun), keep.rownames = "gene")
  }
  clustDE.ls[[myrun]] <- rbindlist(runclust.ls, use.names = TRUE, idcol = "ClusterID")
}
clustDE.dt <- rbindlist(clustDE.ls, use.names = TRUE, idcol = "Expt")
```
### General Plots 

```{r, echo=FALSE}
myso <- expts.integrated.ls$nofilt$SCT.Diff
mymeta <- data.table(myso@meta.data, keep.rownames = "CellID")

# Plot cells/cluster x Expt
clustCts <- mymeta[, .N, by = .(expt.fac, txClusterID.fac)]
clustCts[, txClusterID.fac := droplevels(txClusterID.fac)]
clustCts[, pctCells := N/sum(N)*100, by = expt.fac]
plotcolors <- c("blue", "red", "yellow", "orange")

barplot(data = clustCts, pctCells ~ expt.fac + txClusterID.fac, beside = TRUE, 
        col = plotcolors, horiz = FALSE, las = 1,
        xlab = "Cluster Labels Transferred from DMSO",
        ylab = "% Cells/Expt", cex.names = 0.8,
        main = "Regressed on CC Difference")
legend("topright", legend = levels(clustCts$expt.fac), col = plotcolors, pch = 20)

clustDE.dt[p_val_adj < 0.5, .N, by = .(Expt, ClusterID)]
pval.max <- 0.05
lfc.min <- 0.05

```

### Heatmap of sig genes
Significant genes in combo vs. dmso

```{r, echo=FALSE}
lfc.min <- 0.5
pval.max <- 0.05
min.pct <- 0.1

myso <- expts.integrated.ls$nofilt$SCT.Diff
mymeta <- data.table(myso@meta.data, keep.rownames = "CellID")

for (myclust in allclust) {
  # Create a subset of cells in myclust
  mysubso <- subset(myso, subset = txClusterID.fac == myclust)
  Idents(mysubso) <- "orig.ident"
  
  # Define significant genes
  sigGenes.dt <- clustDE.dt[Expt == "combo" &
                              pct.1 > 0.25 & pct.2 > 0.25 &
                              abs(avg_logFC) > lfc.min &
                              p_val_adj < pval.max &
                              ClusterID == myclust, ]
  
  top10clust <- sigGenes.dt %>% group_by(Expt) %>% top_n(n=25, wt = avg_logFC)

  # Plot significant genes
  IntAssay <- DefaultAssay(mysubso)
  DefaultAssay(mysubso) <- IntAssay
  myplot <- DoHeatmap(mysubso, features = top10clust$gene)
  plot(myplot + plot_annotation(title = sprintf("%s cluster %s, Integrated Object, adjp < %.3f", myregress, myclust, pval.max)))
}
```

```{r, echo=FALSE}
FeaturePlot(expts.integrated.ls$nofilt$SCT.Diff, features = "CD47", assay = "RNA", 
            split.by = "orig.ident")
```
